Pre-processing high frequency animal tracking data

### 2020-08-26 02_filtering_data.Rmd
add move metrics

```{r function_add_outliers}
# function for point outliers
do_add_outliers <- function(data, p_data = 0.1, std_dev = 0.05) {
  n_rows <- floor(p_data * nrow(data))
  tmp_data <- copy(data)

  tmp_data[sample(nrow(data), n_rows),
            c("x", "y") := lapply(.SD,
                                  function(z) {
                                    z + stats::rnorm(n_rows, sd = std_dev)
                                  }),
            .SDcols = c("x", "y")]
  return(tmp_data)
}
```

@@

```{r outlier_error_combinations}
# define unique values
p_outlier <- c(0.01, 0.05, 0.1, 0.25)
ext_outlier <- c(0.01, 0.05, 0.1)
replicate <- seq_len(1)

# make combinations
param_c <- CJ(p_outlier, ext_outlier, replicate)
```

@@

```{r read_landscape}
# read landscape
landscape <- raster("data/data_landscape.tif")

# scale landscape?
bounds <- c(0, 1, 0, 1)
bound_scale <- c(1, 2, 5, 10)
extent(landscape) <- bounds
```

### Make movement track replicates

```{r}
sim_data <- mapply(function(p_outlier, ext_outlier) {
  # make simulated data
  this_data <- do_smoove_data()

  # extract 'true' movement metrics
  speed <- atl_get_speed(this_data, type = "in")

  # extract proportion of landscape used
  land_vals <- extract(landscape,
                       this_data[, c("x", "y")])

  land_vals <- as.data.table(table(landscape))

  # what to return

  # WORK IN PROGRESS HERE

  error_data <- do_add_outliers(this_data, 
                                p_data = p_outlier,
                                std_dev = ext_outlier)
  return(error_data)
}, 
param_c$p_outlier, 
param_c$ext_outlier, 
SIMPLIFY = FALSE)
```

>>>>

```{r function_add_outliers}
# function for point outliers
do_add_outliers <- function(data, p_data = 0.1, std_dev = 0.05) {
  n_rows <- floor(p_data * nrow(data))
  tmp_data <- data.table::copy(data)

  tmp_data[sample(nrow(data), n_rows),
            c("x", "y") := lapply(.SD,
                                  function(z) {
                                    z + stats::rnorm(n_rows, sd = std_dev)
                                  }),
            .SDcols = c("x", "y")]
  return(tmp_data)
}

# function for extracting and tallying landscape values
do_count_landcover <- function(land, track) {
  land_values <- raster::extract(land, track[, c("x", "y")])
  land_values <- data.table::as.data.table(table(land_values))
  land_values[, p_land := N / sum(N)]
  return(land_values)
}
```

@@

```{r outlier_error_combinations}
# define unique values
p_outlier <- c(0.05)
ext_outlier <- c(0.01, 0.05, 0.1)
land_scale <- c(1, 2, 5, 10)
replicate <- seq_len(10)

# make combinations
param_c <- CJ(p_outlier, ext_outlier, land_scale, replicate)
```

@@

```{r read_landscape}
# read landscape
landscape <- raster("data/data_landscape.tif")
```

```{r scale_landscape}
# a function to scale the landscape
do_scale_landscape <- function(land_, scale_factor) {
  bounds <- c(0, 1, 0, 1)
  bounds <- bounds * scale_factor
  raster::extent(land_) <- bounds
  return(land_)
}
```

### Simulate movement over a landscape

```{r}
sim_data <- mapply(function(p_outlier, ext_outlier, scale) {

  # scale the landscape
  tmp_land <- do_scale_landscape(land_ = landscape,
                                 scale_factor = scale)

  # make simulated data
  ref_data <- do_smoove_data()

  # extract 'true' movement metrics
  speed <- atl_get_speed(ref_data, type = "in")
  speed
  land_vals <- do_count_landcover(tmp_land, ref_data)
  land_vals[, type := "reference"]

  # now add error
  error_data <- do_add_outliers(ref_data, 
                                p_data = p_outlier,
                                std_dev = ext_outlier)

  # get movement metrics with error
  error_speed <- atl_get_speed(error_data, type = "in")
  error_land_vals <- do_count_landcover(tmp_land, error_data)
  error_land_vals[, type := "outliers"]

  # now clean data
  clean_data <- copy(error_data)
  clean_data[, `:=`(in_speed = atl_get_speed(clean_data, type = "in"),
                    out_speed = atl_get_speed(clean_data, type = "out"),
                    angle = atl_turning_angle(clean_data))]
  upper_speed_cutoff <- stats::quantile(clean_data$in_speed, 
                                        probs = 0.9,
                                        na.rm = TRUE)
  angle_cutoff <- 40
  clean_data <- atl_filter_covariates(clean_data,
                                      filters = c(glue('{c("in_speed", 
                                                           "out_speed")} < 
                                                         {upper_speed_cutoff}'),
                                                  glue('angle < {angle_cutoff}')
                                                  )
                                      )

  # get speed and land proportion
  clean_speed <- atl_get_speed(clean_data, type = "in")
  clean_land_vals <- do_count_landcover(tmp_land, clean_data)
  clean_land_vals[, type := "cleaned"]

  # synthesize data
  land_props <- rbindlist(list(land_vals, error_land_vals, clean_land_vals))
  # count proportion
  land_props <- dcast(land_props, land_values ~ type, value.var = "p_land")

  land_props[, `:=`(scale = scale,
                    p_outlier = p_outlier,
                    ext_outlier = ext_outlier)]

  # make speed results
  speed_res <- data.table(type = c("reference", "outliers", "cleaned"),
                          speed = c(speed, error_speed, clean_speed))
  results <- list(land_props = land_props,
                  speed_res = speed_res)

  return(land_props)
}, 
param_c$p_outlier, 
param_c$ext_outlier,
param_c$land_scale,
SIMPLIFY = FALSE)
```

```{r}
# get bound list
sim_data <- rbindlist(sim_data)
# get differences between outliers and cleaned wrt ref
#sim_data[, `:=`(d_cleaned = abs(cleaned - reference),
 #               d_outliers = abs(outliers - reference))]

sim_data_long <- copy(sim_data)

sim_data_long <- melt(sim_data_long, 
                      id.vars = c("land_values", "ext_outlier", 
                                  "p_outlier", "scale"))

ggplot(sim_data_long)+
  geom_boxplot(aes(x = factor(land_values),
                   y = value,
                   col = variable))+
  facet_grid(scale ~ ext_outlier, scales = "free")
```
